Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
no message

\subsubsection{Subgroup B Discussion}
The statement ``applied computer science is being attempted in academia without
any formal training'' kicked off our discussion. The group brought expertise in
several different training models, from a two-day Software Carpentry workshop,
a three-week Open Science for Synthesis (OSS) program, to a semester-long
programs. We discussed the pros and cons of different training approaches,
touching on informal learning, for example, where people learning the necessary
skills by asking questions of cross-disciplinary people (``boundary scientists'') in
their work environments.

Some papers explored gaps in training of early-career scientists. Industry
participants in our group confirmed this observation. We asked ourselves, ``Are traditional
courses failing?'' We think yes. Changes to undergraduate curricula requirements
are difficult. But as programming models become more complex, we have to raise
the skill levels. Skills must be improved not only in traditional programming --
learning languages and algorithms -- but also around professional software
development. We want to get to a point where ``Everyone has a new minimum now --
everyone knows Git.'' Developers also need training on licensing choices. Just
because a code is on Git does not mean that it is open.

@@

We asked ourselves, ``Where are there examples of things that are changing
because of this and how can we build momentum?'' One example demonstrates change
over time. In 2007, nanoHUB listed the academic reward structure as a problem in
an EDUCAUSE report where the authors note, ``In the future, nanoHUB researchers
are hoping to change the research culture. While they recognize that young
faculty members are unlikely to get tenure based on their nanoHUB contributions,
they hope to encourage faculty to think beyond their own research needs to
consider publishing tutorials and other content in their fields on the
nanoHUB site.'' Fast forward seven years later where, in a 2014 iSGTW article,
quote nanoHUB Principal Investigator (PI) Gerhard Klimeck, ``A former student of
mine published eight tools on nanoHUB, serving over 6,000 people with his tools.
He then joined a university as a professor and introduced nanoHUB. Use of the
gateway from that university skyrocketed; he used nanoHUB in existing classes,
created new classes, and infused it in his research. Ultimately, the professor's
department head attributed his two-year rise to tenure with the reputation and
innovation he gained through nanoHUB.''

>>>>

\subsubsection{Subgroup B Discussion}
The statement ``applied computer science is being attempted in academia without
any formal training'' kicked off our discussion. The group brought expertise in
several different training models, from a two-day Software Carpentry workshop,
a three-week Open Science for Synthesis (OSS) program, to semester-long
programs. We discussed the pros and cons of different training approaches,
touching on informal learning, for example, where people learn the necessary
skills by asking questions of cross-disciplinary people (``boundary scientists'') in
their work environments.

Some papers explored gaps in training of early-career scientists. Industry
participants in our group confirmed this observation. We asked ourselves, ``Are traditional
courses failing?'' We think yes. Changes to undergraduate curricula requirements
are difficult. But as programming models become more complex, we have to raise
the skill levels. Skills must be improved not only in traditional programming --
learning languages and algorithms -- but also around professional software
development. We want to get to a point where ``Everyone has a new minimum now --
everyone knows Git.'' Developers also need training on licensing choices. Just
because some source code is on Git does not mean that it is open.

@@

We asked ourselves, ``Are there examples of things that are changing
because of this and how can we build momentum?'' One example demonstrates change
over time. In 2007, nanoHUB listed the academic reward structure as a problem in
an EDUCAUSE report where the authors note, ``In the future, nanoHUB researchers
are hoping to change the research culture. While they recognize that young
faculty members are unlikely to get tenure based on their nanoHUB contributions,
they hope to encourage faculty to think beyond their own research needs to
consider publishing tutorials and other content in their fields on the
nanoHUB site.'' Fast forward seven years later where, in a 2014 iSGTW article,
quote nanoHUB Principal Investigator (PI) Gerhard Klimeck, ``A former student of
mine published eight tools on nanoHUB, serving over 6,000 people with his tools.
He then joined a university as a professor and introduced nanoHUB. Use of the
gateway from that university skyrocketed; he used nanoHUB in existing classes,
created new classes, and infused it in his research. Ultimately, the professor's
department head attributed his two-year rise to tenure with the reputation and
innovation he gained through nanoHUB.''
